CRC Follow up: Progress Report 2 - Result Outline

### 2016-10-20 exploratory/notebook/result_outline_20161011.Rmd

#Difference between initial and follow up for thetayc and fit broken down by adenoma and cancer
diff_adn_v_crc <- grid.arrange(
  # Difference from bacterial community structure between adenoma and cancer
  ggplot(difference_table_treatment, aes(factor(dx, levels = c("adenoma", "cancer")), distance, group = 1)) + 
    geom_jitter(aes(color=Dx_Bin), width = 0.3, size = 4, alpha = 0.5) + 
    stat_summary(fun.data = "mean_cl_boot", colour = "black", size = 1) + 
    stat_summary(fun.y = mean, colour = "black", geom = "line") + 
    scale_color_manual(name = "Lesion Type", values = wes_palette("GrandBudapest"), 
                       breaks = c("Adenoma", "adv Adenoma", "Cancer"), labels = c("Adenoma", "SRN", "Cancer")) + 
    coord_cartesian(ylim = c(0, 1)) + ylab("Thetayc Distance") + xlab("") + theme_bw() + 
    theme(axis.title = element_text(face="bold"), legend.title = element_text(face="bold"), 
          title = element_text(face="bold")) + 
    annotate("text", label = paste("P-value = ", round(pValueList[["thetaDiff"]], digits = 2)), x = 1.5, y = 0.75),

  # Difference from fit between adenoma and cancer
  ggplot(difference_table_treatment, aes(factor(dx, levels = c("adenoma", "cancer")), fit_difference, group = 1)) + 
    geom_jitter(aes(color=Dx_Bin), width = 0.3, size = 4, alpha = 0.5) + 
    stat_summary(fun.data = "mean_cl_boot", colour = "black", size = 1) + 
    stat_summary(fun.y = mean, colour = "black", geom = "line") + 
    scale_color_manual(name = "Lesion Type", values = wes_palette("GrandBudapest"), 
                       breaks = c("Adenoma", "adv Adenoma", "Cancer"), labels = c("Adenoma", "SRN", "Cancer")) + 
    ylab("Change in Fit from Initial to Follow Up") + xlab("") + theme_bw() + 
    theme(axis.title = element_text(face="bold"), legend.title = element_text(face="bold"), 
          title = element_text(face="bold")) + 
    annotate("text", label = paste("P-value = ", format(round(pValueList[["fitDiff"]], digits = 7))), x = 1.5, y = 0)
)

@@

# Load in necessary Data
testSetAUC <- read.csv("results/tables/testSetAUC.csv", stringsAsFactors = F, header = T)
rownames(testSetAUC) <- testSetAUC$V1
colnames(testSetAUC) <- c("Model", "AUC")
wFit_ROC_table <- read.csv("results/tables/ROCCurve_sens_spec.csv")
woFit_ROC_table <- read.csv("results/tables/ROCCurve_noFit_sens_spec.csv")

@@

@@

graphData <- rbind(
  mutate(wFit_ROC_table, FIT = rep("wFit", length(rownames(wFit_ROC_table)))) %>% 
    filter(model == "lesionALL"), 
  mutate(woFit_ROC_table, FIT = rep("woFit", length(rownames(woFit_ROC_table)))) %>% 
    filter(model == "lesionALL")
)

# Get Pvalue difference between two ROCs

ggplot(graphData, aes(sensitivities, specificities)) + 
  geom_line(aes(group = FIT, color = FIT), size = 1.5) + scale_x_continuous(trans = "reverse") + 
  theme_bw() + xlab("Sensitivity") + ylab("Specificity") + geom_abline(intercept = 1, linetype = 2) + 
  scale_color_discrete(name = "Lesion Model", breaks = c("wFit", "woFit"), 
                       labels = c(paste("with Fit\n", "(Test Set AUC = ", 
                                        round(testSetAUC['withFit', 'AUC'], digits = 3), ")", sep=""), 
                                  paste("without Fit\n", "(Test Set AUC = ", 
                                        round(testSetAUC['woFit', 'AUC'], digits = 3), ")", sep=""))) + 
  theme(axis.title = element_text(face = "bold"), 
        legend.position = c(0.75, 0.4), legend.title = element_text(face="bold")) + 
  annotate("text", label = paste("P-value = ", round(roc.test(wFit_lesion, woFit_lesion)$p.value, 
                                                     digits = 3)), x = 0.9, y = 0.9, size = 4)

@@

**Figure 3:** ROC curve of the lesion model with and without Fit.  Overall there is no significant difference between the two models.  This was built using a total of ```r length(rownames(metaI))``` individuals which was then tested on ```r length(rownames(metaF))``` which had initial and follow up data available.

@@

lesion_selected_taxa <- tax_df[filter(wofit_pvalue_summary, OTU_WoF_adjust_pvalues < 0.05)[, 'otus'], ]
lesion_selected_labs <- createTaxaLabeller(lesion_selected_taxa)

cancer_positive <- as.data.frame(cbind(rep(as.character(good_metaf$Disease_Free), 2), rep(good_metaf$EDRN, 2))) %>% 
  rename(Disease_Free = V1, EDRN = V2)

>>>>

#Difference between initial and follow up for thetayc and fit broken down by adenoma and cancer
diff_adn_v_crc <- grid.arrange(
  # Difference from bacterial community structure between adenoma and cancer
  ggplot(difference_table_treatment, aes(factor(dx, levels = c("adenoma", "cancer")), distance, group = 1)) + 
    geom_jitter(aes(color=Dx_Bin), width = 0.3, size = 4, alpha = 0.5) + 
    stat_summary(fun.data = "mean_cl_boot", colour = "black", size = 1) + 
    stat_summary(fun.y = mean, colour = "black", geom = "line") + 
    scale_color_manual(name = "Lesion Type", values = wes_palette("GrandBudapest"), 
                       breaks = c("Adenoma", "adv Adenoma", "Cancer"), labels = c("Adenoma", "SRN", "Cancer")) + 
    coord_cartesian(ylim = c(0, 1)) + ylab("Thetayc Distance") + xlab("") + theme_bw() + 
    theme(axis.title = element_text(face="bold"), legend.title = element_text(face="bold"), 
          title = element_text(face="bold")) + 
    annotate("text", label = paste("P-value = ", round(pValueList[["thetaDiff"]], digits = 2)), x = 1.5, y = 0.7),

  # Difference from fit between adenoma and cancer
  ggplot(difference_table_treatment, aes(factor(dx, levels = c("adenoma", "cancer")), fit_difference*-1, group = 1)) + 
    geom_jitter(aes(color=Dx_Bin), width = 0.3, size = 4, alpha = 0.5) + 
    stat_summary(fun.data = "mean_cl_boot", colour = "black", size = 1) + 
    stat_summary(fun.y = mean, colour = "black", geom = "line") + 
    scale_color_manual(name = "Lesion Type", values = wes_palette("GrandBudapest"), 
                       breaks = c("Adenoma", "adv Adenoma", "Cancer"), labels = c("Adenoma", "SRN", "Cancer")) + 
    ylab("Change in Fit from Follow up to Initial") + xlab("") + theme_bw() + 
    theme(axis.title = element_text(face="bold"), legend.title = element_text(face="bold"), 
          title = element_text(face="bold")) + 
    annotate("text", label = paste("P-value = ", format(round(pValueList[["fitDiff"]], digits = 7))), x = 1.5, y = 100)
)

@@

# Load in necessary Data
testSetAUC <- read.csv("results/tables/testSetAUC.csv", stringsAsFactors = F, header = T)
rownames(testSetAUC) <- testSetAUC$V1
colnames(testSetAUC) <- c("Model", "AUC")
wFit_ROC_table <- read.csv("results/tables/ROCCurve_sens_spec.csv")
woFit_ROC_table <- read.csv("results/tables/ROCCurve_noFit_sens_spec.csv")
wFit_test_ROC_table <- read.csv("results/tables/wFit_test_ROCCurve_sens_spec.csv")
woFit_test_ROC_table <- read.csv("results/tables/woFit_test_ROCCurve_sens_spec.csv")

@@

load("exploratory/wFit_Test_ROC.RData")
test_wFit_lesion <- lesion_combined_ROC
rm(lesion_combined_ROC)

load("exploratory/woFit_Test_ROC.RData")
test_woFit_lesion <- lesion_combined_ROC
rm(lesion_combined_ROC)

@@

graphData <- rbind(
  mutate(wFit_ROC_table, FIT = rep("wFit", length(rownames(wFit_ROC_table)))) %>% 
    filter(model == "lesionALL"), 
  mutate(woFit_ROC_table, FIT = rep("woFit", length(rownames(woFit_ROC_table)))) %>% 
    filter(model == "lesionALL"), 
  mutate(wFit_test_ROC_table, FIT = rep("wFitTest", length(rownames(wFit_test_ROC_table)))), 
  mutate(woFit_test_ROC_table, FIT = rep("woFitTest", length(rownames(woFit_test_ROC_table))))
)

#Get the pvalues

corr_pvalue_ROC_table <- getROCPvalue(rocNameList = list(
  wfit = wFit_lesion, wofit = woFit_lesion, test_wfit = test_wFit_lesion, test_wofit = test_woFit_lesion), 
  modelList = c("wfit", "wofit", "test_wfit", "test_wofit"), 4, multi = F)

# Get Pvalue difference between two ROCs (Plot the test set data as well)

ggplot(graphData, aes(sensitivities, specificities)) + 
  geom_line(aes(group = FIT, color = FIT), size = 1.5) + scale_x_continuous(trans = "reverse") + 
  theme_bw() + xlab("Sensitivity") + ylab("Specificity") + geom_abline(intercept = 1, linetype = 2) + 
  scale_color_discrete(name = "Lesion Model", breaks = c("wFit", "woFit", "wFitTest", "woFitTest"), 
                       labels = c(paste("with Fit (Train Set)\n", sep = ""), 
                                  paste("without Fit (Train Set)\n", sep=""), 
                                  paste("with Fit (Test Set)\n", sep=""), 
                                  paste("without Fit (Test Set)\n", sep=""))) + 
  theme(axis.title = element_text(face = "bold"), 
        legend.position = c(0.75, 0.4), legend.title = element_text(face="bold"))

@@

**Figure 3:** ROC curve of the lesion model with and without Fit for the training and test set.  Overall there was no significant difference between with and without Fit models for the training set (```r paste("P-value = ", round(roc.test(wFit_lesion, woFit_lesion)$p.value, digits = 3))```).  This was built using a total of ```r length(rownames(metaI))``` individuals which was then tested on ```r length(rownames(metaF))``` which had initial and follow up data available.

@@

lesion_selected_taxa <- tax_df[filter(wofit_pvalue_summary, OTU_WoF_adjust_pvalues < 0.05 | otus == "Otu000205")[, 'otus'], ]
lesion_selected_taxa <- lesion_selected_taxa[c("Otu000126", "Otu000397", "Otu000205"), ]
lesion_selected_labs <- createTaxaLabeller(lesion_selected_taxa)

cancer_positive <- as.data.frame(cbind(rep(as.character(good_metaf$Disease_Free), 3), rep(good_metaf$EDRN, 3))) %>% 
  rename(Disease_Free = V1, EDRN = V2)
